How to compute the fusion product of MV cycles in type A

### 2021-03-16 draft.tex
joel changed some notation and finished the proof of theorem 3

    We also have a lattice description 
    $$ 
    \Gr = \left\{ L \subset \cK^m : \text{ $L$ is a $\cO$-lattice} \right\}\,.
    $$
    We obtain the lattice from $ (V,\varphi) $ by setting $ L = \Gamma(D_0, V)$ which is embedded into $ \cK^m = \Gamma(D_0^\times, V_0)$ using $ \varphi$.
    % 
    \item For any $ s \in \CC $, 
    % \acom{should it be $\AA$ instead of $\CC$} 
    the ordinary affine Grassmannian $\Gr_s = G(\cK_s)/G(\cO_s)$ at $ s $. It is described in modular terms by
    $$
    \Gr_s = \left\{ (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_s $, $\varphi : V \xrightarrow{\sim} V_0 $ on $ D^\times_s $} \right\}\,.
    $$
    % where $D_s$ is the formal disc centered at $s$ viewed as an element of $\PP$.
    % \acom{What about fixing the notation $V_0$ somewhere?}

@@

@@

Next, take $p$ to be a permutation of the (lengths of) columns $a_j$ of $\lambda$ such that $a_{p(1)},\dots,a_{p(m_1)}$ are the (lengths of) columns of $\lambda_1$ and $a_{p(m_1 + 1)},\dots, a_{p(m_1 + m_2)}$ are the (lengths of) columns of $\lambda_2$. 
Then the functions 
$$
r(1,j) = N - a_{p(1)} - \cdots - a_{p(j)} \text{ and } r(2,j) = N - a_{p(m_1 + 1)} - \cdots - a_{p(m_1 + j)}
$$
are such that 
\begin{align*}
    \OO^{\lambda_1,\lambda_2} = X_{r,(0,s)} = \{A:\rk A^j &= r(1,j) \text{ for } i = 1,\dots,\lambda_{1,1} \\ \text{ and } \rk (A-s)^j &= r(2,j) \text{ for } i = \lambda_{1,1} + 1 ,\dots, \lambda_{2,1} \}
\end{align*}
and according to \cite[Theorem 2.1(iii)]{eisenbud1989rank} the family $X_{r,\AA^m}$ is flat over $\AA^m$ with zero fibre $X_{r,0} = \OO^\lambda$ as desired.

@@

    Now assume that $ (L, s) \in \Gr^{\lambda_1, \lambda_2}_{0,s}$. We know that \acom{by definition?} $ L(0) \in \Gr^{\lambda_1}$ and $L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on 
    % $ \CC\xt^m / L(0)$ 
    $\cO^m/L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on 
    % $ \CC\xt[t-s]^m / L(s) $ 
    $\cO_s^m/L(s)$ has Jordan type $ \lambda_2$.  Now, the map 
    % $ L_0 / L \rightarrow \CC\xt^m / L(0)$ 
    $L_0/L\to \cO^m/L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \cO^m / L(0)$.   The same thing holds for the $s$-generalized eigenspace of $t $ and $ \cO_s^m/L(s)$. This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i).

    On the other hand, if $ L = g t^{\lambda_1} (t-s)^{\lambda_2} $ for some $ g \in G(\CC[t])$, then $ L(0) = g t^{\lambda_1} $ since 
    % $ (t-s)^{\lambda_2} \in G(\CC\xt)$. 
    $ (t-s)^{\lambda_2} \in G(\cO)$. In this way, we see that (iii) implies (ii) and the logic can be reversed to get the equivalence.
\end{proof}

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
                \item $ L = \Sp_\cO(v_1, \dots, v_m)$ for some $ v_i $ of the form $ v_i = t^{\mu_i} e_i + \sum_{j=1}^m p_{ij} e_j $ where $ p_{ij} \in \CC[t] $ has degree less than $ \min(\mu_i, \mu_j)$. \acom{Do we want to sum over $j \ne i$ here? Maybe nvm because in $\TT_\mu$ the diagonal blocks can have $*$'s in the last row. Also $\cO$ should be $\CC[t]$.}
        \item  For all $ i $, 
        $$ t^{\mu_i} e_i \in \Sp_\CC(\{t^k e_j : 0 \le k < \min(\mu_i, \mu_j), 1 \le j \le m \}) + L. $$
    \end{enumerate}
    Moreover, for such $L $, $ \beta_\mu := \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$.
\end{lemma}

@@

\begin{proof}
    Let $ L \in \cW_\mu$.  Then $ L = \Sp_\cO(v_1, \dots, v_m) $ for some $ v_i $ with $ v_i = t^{\mu_i} e_i + \sum_{j=1} q_{ij}t^{\mu_i} e_j $ and $ q_{ij} \in t^{-1} \Oinf$.  Since $ L \in \Grth^+ $, we see that $ v_i \in L_0$ which means that $ p_{ij} := q_{ij}t^{\mu_i} $ lies in $ \CC[t]$.  By construction, the polynomial $ p_{ij}$ has degree less than $ \mu_i$.

@@

    Suppose that $ L = \Sp_\cO(v_1, \dots, v_m)$ as in (ii).  Then $[t^{\mu_i} e_i] \in \Sp_\CC(\beta_\mu) $ in $ L_0/L$ and we see that 
    $$t^{\mu_i} e_i \in \Sp_\CC(\{t^k e_j : 0 \le k < \mu_i \}) + L\,. $$
    Hence (ii) implies (iii).

@@

    Finally to show that $ \beta_\mu$ forms a basis for $ L_0/L$, it suffices to show that for each $ i$, $ t^{\mu_i} e_i  \in \Sp_\CC(\beta_\mu) + L$.  This follows immediately from (iii).
    % 
    \end{proof}

@@

\begin{proof}
    Let $ L $ lie in this image.  Then $ L = g t^{\mu_1} (t-s)^{\mu_2}$ fom some $ g \in N(\CC[t, t^{-1}, (t-s)^{-1}]) $.  Let $ h =(t-s)^{\mu_2} t^{-\mu_2}  $.  Note that $ h \in T_1(\Oinf)$. 
    Moreover,
    $$ L = h (h^{-1} g h) t^{\mu}\,. $$
    Note that 
    % $ h^{-1} g h \in N(\CC\xT[t^{-1}])$ 
    $h^{-1} g h \in N(\cK_\infty)$ and from \cite{kamnitzer2014yangians} 
    \jcom{Check this}, we know that $ (h^{-1} g h) t^{\mu} \in \cW_\mu$ and so the result follows.

    \acom{
        % The argument from my thesis does not go through since $\cW_\mu$ is defined mod $\CC[t]$. More precisely, we still have $N(\Kinf) = N_1(\CC[t]) N(\CC\xt[t^{-1}])$ and so $h^{-1}g h=g_1g_2$ for some $g_1 \in N_1(\CC[t])$ for some $g_2\in N\xt[t^{-1}]$. But now we would like to write $g_2 t^\mu = t^\mu g_3$ for some $g_3\in\CC[t]$. Oh, but I guess we can also 
        To check this write $N(\Kinf) = N_1(\Oinf)N(\CC[t])$. Does it hold? If $\mu$ is dominant and $g\in N(\CC[t])$ then $gt^\mu = t^\mu g'$ for some $g'\in N(\CC[t])$.} % What is a textbook reference for these factorization claims? are they obvious; they seem obvi
        \jcom{Yes, that is correct.  I think that I once had Alex write out a careful proof of this fact.  It holds for any unipotent group.  The proof is by induction.}
    \end{proof}

@@

\begin{theorem} \label{th:TmuWmu}
The map $ \TT_\mu \rightarrow \Grth^+ \cap \cW_\mu $ given by $ A \mapsto g(A) $ is an isomorphism with inverse given by
$$ L \mapsto [t|_{L_0/L} ]_{\beta_\mu} \,. $$
\acom{Up to transpose?}
\end{theorem}

@@

\begin{theorem} \label{th:OGrl}
    There is an isomorphism
    $$\overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap \TT_\mu \cong \overline\Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
    given by $ (A,s) \mapsto (g(A), s)$.
\end{theorem}
\acom{Joel, you stopped writing $X_{0,\AA}$ in favour of $X_\AA$ --- should we do this everywhere?}
\jcom{Oops, no I think that we should stick with $ X_{0, \AA}$.}
\begin{proof}
Since we already have the isomorphism from Theorem \ref{th:TmuWmu}, it suffices to show that for any $ A \in \TT_\mu$, $$ (A,s) \in \overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap \TT_\mu \text{ if and only if } (g(A), s) \in \overline\Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
This follows immediately from Lemma \ref{le:Grl1l2}.
\end{proof}

>>>>

    We also have a lattice description 
    $$ 
    \Gr = \left\{ L \subset \cK^m : \text{ $L$ is a $\cO$-lattice} \right\}\,.
    $$
    We obtain the lattice from $ (V,\varphi) $ by setting $ L = \Gamma(D_0, V)$ which is embedded into $ \cK^m = \Gamma(D_0^\times, V_0)$ using $ \varphi$.  On the other hand, we obtain the lattice from the group-theoretic description $ G(\cK) / G(\cO) $ by setting $ L = g \cO^m$ for $ g \in G(\cK)$.
    % 
    \item For any $ s \in \CC $, 
    % \acom{should it be $\AA$ instead of $\CC$} 
    the ordinary affine Grassmannian $\Gr_s = G(\cK_s)/G(\cO_s)$ at $ s $. It is described in modular terms by
    $$
    \Gr_s = \left\{ (V, \varphi) : \text{$V$ is a rank $m$ vector bundle on $ D_s $, $\varphi : V \xrightarrow{\sim} V_0 $ on $ D^\times_s $} \right\}\,.
    $$
    % where $D_s$ is the formal disc centered at $s$ viewed as an element of $\PP$.
    % \acom{What about fixing the notation $V_0$ somewhere?}

@@

    In group-theoretic terms, this isomorphism is given using the diagonal map
    $$G(\CC[t, t^{-1}, (t-s)^{-1}])/G(\CC[t]) \rightarrow G(\cK)/G(\cO) \times G(\cK_s)/G(\cO_s)
    $$
    This is compatible with the lattice description since for $ g \in G(\CC[t, t^{-1}, (t-s)^{-1}])/G(\CC[t]) $, we have
    $$
    g \CC[t]^m \otimes_{\CC[t]} \cO = g \cO^m
    $$
    and similarly for $ L(s)$.

@@

Next, take $p$ to be a permutation of the (lengths of) columns $a_j$ of $\lambda$ such that $a_{p(1)},\dots,a_{p(m_1)}$ are the (lengths of) columns of $\lambda_1$ and $a_{p(m_1 + 1)},\dots, a_{p(m_1 + m_2)}$ are the (lengths of) columns of $\lambda_2$. 
Then the functions 
$$
\begin{aligned}
r(1,j) &= N - a_{p(1)} - \cdots - a_{p(j)} \\
% \text{ and } 
r(2,j) &= N - a_{p(m_1 + 1)} - \cdots - a_{p(m_1 + j)}
\end{aligned}
$$
are such that 
\begin{align*}
    \OO^{\lambda_1,\lambda_2}_{0,s} = X_{r,(0,s)} = \{A:\rk A^j &= r(1,j) \text{ for } i = 1,\dots, m_1 \\ \text{ and } \rk (A-s)^j &= r(2,j) \text{ for } i = m_1 + 1 ,\dots, m \}
\end{align*}
and according to \cite[Theorem 2.1(iii)]{eisenbud1989rank} the family $X_{r,\AA^m}$ is flat over $\AA^m$ with zero fibre $X_{r,0} = \OO^\lambda$ as desired. I guess we need to carefully choose the $W \subset \AA^m$ defining our family, which is smaller than the family $X_{r,\AA^m}$. Well, $W$ is just cut out by $0 = x_1 = x_2 = \cdots = x_{m_1}$ and $ x_{m_1 + 1} = \cdots = x_m$. Or some permutation thereof to match the permutation $p$ above.

@@

    Now assume that $ (L, s) \in \Gr^{\lambda_1, \lambda_2}_{0,s}$. We know that \acom{by definition?} $ L(0) \in \Gr^{\lambda_1}$ and $L(s) \in \Gr^{\lambda_2} $.  This means that $t $ acting on 
    % $ \CC\xt^m / L(0)$ 
    $\cO^m/L(0)$ has Jordan type $ \lambda_1$ and $ t$ acting on 
    % $ \CC\xt[t-s]^m / L(s) $ 
    $\cO_s^m/L(s)$ has Jordan type $ \lambda_2$.  Now, the map 
    % $ L_0 / L \rightarrow \CC\xt^m / L(0)$ 
    $\CC[t]^m/L\to \cO^m/L(0)$ induces an isomorphism between the $0$-generalized eigenspace of $ t$ and $ \cO^m / L(0)$.   The same thing holds for the $s$-generalized eigenspace of $t $ and $ \cO_s^m/L(s)$. This shows that (i) implies (ii) and the logic can be reversed to see that (ii) implies (i).

    On the other hand, if $ L = g t^{\lambda_1} (t-s)^{\lambda_2} \CC[t]^m$ for some $ g \in G(\CC[t])$, then $ L(0) = g t^{\lambda_1}\cO^m $ since 
    % $ (t-s)^{\lambda_2} \in G(\CC\xt)$. 
    $ (t-s)^{\lambda_2} \in G(\cO)$. In this way, we see that (iii) implies (ii) and the logic can be reversed to get the equivalence.
\end{proof}

\begin{lemma} \label{le:Wmu}
    Let $ L \in \Grth^+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in \cW_\mu$.
                \item $ L = \Sp_{\CC[t]}(v_1, \dots, v_m)$ for some $ v_i $ of the form $ v_i = t^{\mu_i} e_i + \sum_{j=1}^m p_{ij} e_j $ where $ p_{ij} \in \CC[t] $ has degree less than $ \min(\mu_i, \mu_j)$. \acom{Do we want to sum over $j \ne i$ here? Maybe nvm because in $\TT_\mu$ the diagonal blocks can have $*$'s in the last row. Also $\cO$ should be $\CC[t]$.}
        \item  For all $ i $, 
        $$ t^{\mu_i} e_i \in \Sp_\CC(\{t^k e_j : 0 \le k < \min(\mu_i, \mu_j), 1 \le j \le m \}) + L. $$
    \end{enumerate}
    Moreover, for such $L $, $ \beta_\mu := \{ [t^k e_i] : 0 \le k < \mu_i, 1 \le i \le m\}$ forms a basis for $ L_0/L$.
\end{lemma}

@@

\begin{proof}
    Let $ L \in \cW_\mu$.  Then $ L = \Sp_{\CC[t]}(v_1, \dots, v_m) $ for some $ v_i $ with $ v_i = t^{\mu_i} e_i + \sum_{j=1} q_{ij}t^{\mu_i} e_j $ and $ q_{ij} \in t^{-1} \Oinf$.  Since $ L \in \Grth^+ $, we see that $ v_i \in L_0$ which means that $ p_{ij} := q_{ij}t^{\mu_i} $ lies in $ \CC[t]$.  By construction, the polynomial $ p_{ij}$ has degree less than $ \mu_i$.

@@

    Suppose that $ L = \Sp_{\CC[t]}(v_1, \dots, v_m)$ as in (ii).  Then
    $$t^{\mu_i} e_i - v_i \in \Sp_\CC(\{t^k e_j :  k < \min(\mu_i, \mu_j), 1 \le j \le m \})  \,. $$
    Hence (ii) implies (iii).

@@

    Finally to show that $ \beta_\mu$ forms a basis for $ \CC[t]^m/L$, it suffices to show that for each $ i$, $ t^{\mu_i} e_i  \in \Sp_\CC(\beta_\mu) + L$.  This follows immediately from (iii).
    % 
    \end{proof}

@@

\begin{proof}
    Let $ L $ lie in this image.  Then $ L = g t^{\mu_1} (t-s)^{\mu_2}\CC[t]^m$ for some $ g \in N(\CC[t, t^{-1}, (t-s)^{-1}]) $.  Let $ h =(t-s)^{\mu_2} t^{-\mu_2}  $.  Note that $ h \in T_1(\Oinf)$. 
    Moreover,
    $$ L = h (h^{-1} g h) t^{\mu}\CC[t]^m\,. $$
    Note that 
    % $ h^{-1} g h \in N(\CC\xT[t^{-1}])$ 
    $h^{-1} g h \in N(\cK_\infty)$ and so we can factor $ h^{-1} g h = n_1 n_2$ for $ n_1 \in N_1(\CC\xt[t^{-1}]), n_2 \in N(\CC[t]) $.  Since $ \mu $ is dominant, we see that $ t^{-\mu} n t^\mu \in N(\CC[t]) $.  Thus, $ L = h n_1 t^\mu \CC[t]^m$.  Since $ hn_1 \in G_1(\CC\xt[t^{-1}])$, the result follows.
    \end{proof}

@@

\begin{theorem} \label{th:TmuWmu}
The map $ \TT_\mu \rightarrow \Grth^+ \cap \cW_\mu $ given by $ A \mapsto g(A)\CC[t]^m $ is an isomorphism with inverse given by
$$ L \mapsto [t|_{\CC[t]^m/L} ]_{\beta_\mu} \,. $$
\acom{Up to transpose?}
\end{theorem}

@@

\begin{theorem} \label{th:OGrl}
    There is an isomorphism
    $$\overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap \TT_\mu \cong \overline\Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
    given by $ (A,s) \mapsto (g(A)\CC[t]^m, s)$.
\end{theorem}
\acom{Joel, you stopped writing $X_{0,\AA}$ in favour of $X_\AA$ --- should we do this everywhere?}
\jcom{Oops, no I think that we should stick with $ X_{0, \AA}$.}
\begin{proof}
Since we already have the isomorphism from Theorem \ref{th:TmuWmu}, it suffices to show that for any $ A \in \TT_\mu$, $$ (A,s) \in \overline{\OO}^{\lambda_1, \lambda_2}_\AA \cap \TT_\mu \text{ if and only if } (g(A)\CC[t]^m, s) \in \overline\Gr^{\lambda_1, \lambda_2}_\AA \cap \cW_\mu $$
This follows immediately from Lemma \ref{le:Grl1l2}.
\end{proof}
